---
title: Trump's Twitter Ratio this year and his most unpopular tweets
author: Rajesh Korde
date: '2017-10-20'
slug: trump-s-twitter-ratio-this-year-and-his-most-unpopular-tweets
categories:
  - Analysis
tags:
  - politics
  - R
  - twitter
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, 
                      message=FALSE, cache=TRUE)
```

```{r}
library(feather)
library(tidyverse)
library(stringr)
library(lubridate)
```

[Twitter Ratio](http://knowyourmeme.com/memes/the-ratio) is an unofficial measure of how unpopular a particular tweet is. It is calculated simply by dividing the number of replies to the tweet by the number of favorites (or retweets). The logic being if its a crappy tweet, you would get fewer favorites and more replies. Thus higher the ratio, worse the tweet. In this post, I look at Trump's twitter ratio since the day of his inauguration and his most unpopular tweets (as in tweets with highest twitter ratio).

<!--more-->

I calculate twitter ratio as replies divided by favorites and not retweets, because I have found many people ironically retweet Trump's tweets. But I doubt anyone favorites his tweets ironically. Given below is his twitter ratio calculated weekly. The trend line shows that his twitter ratio has been gradually increasing at a slow rate. However, this may be because his twitter ratio was abnormally low in the first couple of weeks of his presidency. What I found more striking is the wild variations in the ratio (changes by almost 2x). I think its reflective of the chaotic nature of his presidency where a storm is always brewing anytime there is some quiet.


```{r}

tab <- read_feather(path = "D:/GitRepos/Github/RTestCode/Data/trumptwitter.feather")

invisible_tweets <- filter(tab, grepl(pattern = "private lunch|LIMITED EDITION", x = text)) %>%
  select(id_str)

ratio_tweets <- filter(tab, replies > 50 & favorites > 50) %>%
  filter(!id_str %in% invisible_tweets$id_str) %>%
  mutate(ratio = replies/favorites) %>%
  arrange(desc(ratio))

```

```{r}
library(stringr)
library(lubridate)

month_lookup <- 1:12
names(month_lookup) <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                         "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

fix_date <- function(dt) {
  y <- str_split(dt, " ")[[1]][c(2, 3, 6)]
  y[1] <- unname(month_lookup[y[1]])
  return(mdy(paste(y, collapse = " ")))
}

d <- ratio_tweets %>% 
  rowwise() %>%
  bind_cols(do(., {
    data.frame(date = fix_date(.$created_at))
  })) %>%
  ungroup() %>%
  mutate(week = week(date))


by_week <- d %>%
  filter(date > mdy("1/20/2017")) %>%
  group_by(week) %>%
  summarise(replies = sum(replies), favorites = sum(favorites)) %>%
  mutate(ratio = replies/favorites) %>%
  mutate(date = ymd("2017-01-01") + weeks(week))


ggplot(by_week, aes(date, ratio)) + 
  geom_line() +
  geom_smooth(se = FALSE, method = "lm") +
  xlab("") +
  ylab("Twitter Ratio") +
  ggtitle("Trump's weekly twitter ratio") +
  theme_bw()


```


Given below are the top 25 tweets with the worst twitter ratio.

```{r}
library(kableExtra)
options(scipen=1, digits=2)

top <- d %>%
  mutate(tweet_text = paste0("[", text, "](", url, ")")) %>%
  mutate(tweet_text = str_replace(tweet_text, "@", "")) %>%
  select(Tweet = tweet_text, 
         Created = date, 
         Replies = replies,
         Retweets = retweets,
         Favorites = favorites,
         Ratio = ratio) %>%
  slice(1:25)

kable(top) %>%
  kable_styling(full_width = F)
```

I got the data for Trump's tweets from [trumptwitterarchive.com](http://www.trumptwitterarchive.com/). You can find all the source code for this blog post [here](https://github.com/rajkorde/RTestCode/blob/master/Scripts/TrumpTwitterArchive.R).